---
title: "Homework #2"
author: "Kobe Pranivong; illuminobe"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## R commands
1. Today's temperature has a high of 24 and a low of 15 degree fahrenheit. Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
fahr <- c(24, 15)
cel <- c()
for (temp in c(24, 15)) {
  cel <- c(cel, (temp - 32) * 5/9)
}
cel
```

2. Create a vector of length 200. The odd entries should be 3 and even entries should be 1. Storage the created in a variable named `vec200` and print.
```{r}
vec200 <- c()
for (x in c(1:200)) {
  if (x %% 2 == 0) {
    vec200 <- c(vec200, 1)
  }
  else {
    vec200 <- c(vec200, 3)
  }
}
vec200
```

3. How many perfect squares are there between 1 and 2001? Storage the answer in a variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
numPerfect = floor(sqrt(2001))
numPerfect
```
There are 44 perfect squares between 1 and 2001.

## Cars and mileage
1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.
```{r}
library(ggplot2)
top3 = mpg[order(mpg$hwy, decreasing = TRUE), ][1:3, ]
top3
```
The 3 most fuel-efficient models in terms of highway mileage are jetta, new beetle, and new beetle.

2. How many different compact models are included in this dataset? Storage the answer in a variable named `numCompact` and print.
```{r}
models = mpg$model[mpg$class == "compact"]
numCompact = unique(models)
numCompact
```
There are 8 different compact models included in this dataset.

3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible? Explain your findings.
```{r}
plot(mpg$hwy, mpg$cty, main = "Highway vs City MPG", xlab = "Highway MPG", ylab = "City MPG")
```

There is a positive direct linear relationship between highway and city fuel-efficiency. There are so few points visible because they are overlapping.

4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. Explain your findings.
```{r}
plot(density(mpg$hwy[mpg$year == 1999]), col = "red", main="Highway MPG")
lines(density(mpg$hwy[mpg$year == 2008]), col = "blue")
legend("right", legend=c("1999", "2008"), col=c("red", "blue"), lty=c(1,1))
```
```{r}
plot(density(mpg$cty[mpg$year == 1999]), col = "red", main="City MPG")
lines(density(mpg$cty[mpg$year == 2008]), col = "blue")
legend("right", legend=c("1999", "2008"), col=c("red", "blue"), lty=c(1,1))
```
```{r}
hwyBox <- ggplot(mpg, aes(group = year, x = year, y = hwy)) + geom_boxplot()
hwyBox
ctyBox <- ggplot(mpg, aes(group = year, x = year, y = cty)) + geom_boxplot()
ctyBox
```
```{r}
city_data_1999 = mpg$cty[mpg$year == 1999]
summary(city_data_1999)
city_data_2008 = mpg$cty[mpg$year == 2008]
summary(city_data_2008)
```
```{r}
hwy_data_1999 = mpg$hwy[mpg$year == 1999]
summary(hwy_data_1999)
hwy_data_2008 = mpg$hwy[mpg$year == 2008]
summary(hwy_data_2008)
```

My conclusion is that cars in 2008 are not more fuel efficient than cars in 1999. 

Based on the data from the summary(), when it comes to driving in the city, both years had a median of 17.0 mpg. Additionally, cars in 1999 had a mean of 17.02 mpg while cars in 2008 had a mean of 16.7 mpg. When it comes to driving on the highway, cars in 1999 had a median of 25.00 mpg while cars in 2008 had a median of 24.00 mpg. Additionally, cars in 1999 had a mean of 23.43 mpg while cars in 2008 had a mean of 23.45 mpg. These numbers show that the fuel efficiency is comparable for cars in 1999 and 2008.

Based on the graphs, both the boxplots and the line graphs show that the fuel efficiency is comparable for cars in 1999 and 2008.